Supplementary Materials

### 2023-08-26 reports/state_figures.Rmd
update figures

state_res %>%
  select(-c(date)) %>%
  distinct() %>%
  mutate(interval_length =exp_cases_ub - exp_cases_lb) %>%
  group_by(vlabel) %>%
  summarize(mean_interval_length = mean(interval_length))

@@

>>>>

state_res %>%
  select(-c(date)) %>%
  distinct() %>%
  mutate(interval_length =exp_cases_ub - exp_cases_lb) %>%
  group_by(vlabel) %>%
  summarize(mean_interval_length = mean(interval_length)) %>%
  arrange(desc(mean_interval_length))

state_res_differenced <- state_res %>%
  select(-date) %>% distinct() %>%
  group_by(state, vlabel) %>%
  mutate(exp_cases_median_differenced = exp_cases_median - lag(exp_cases_median, n =1, order_by = biweek),
         infections_differenced = infections - lag(infections, n = 1, order_by=biweek)) %>%
  select(state, name, vlabel, contains('differenced'), biweek) %>%
  filter(!is.na(exp_cases_median_differenced)) %>%
  filter(biweek <=25)

# if lag is NEGATIVE, covidestim LEADS wastewater
# if lag is POSITIVE, covidestim LAGS wastewater
# since by the documentation, 'the lag k value returned by ccf(x, y) estimates
# the correlation between x[t+k] and y[t].'

get_max_lag_by_version <- function(input_df_for_version,
                                   varnames = c("infections_differenced",
                                  "exp_cases_median_differenced"),
                                  comparison = "bias corrected") {

  var1 <- varnames[1]
  var2 <-varnames[2]
  lab <- unique(input_df_for_version$vlabel)

  cross_correlations <- ccf(
    pull(input_df_for_version, var1),
     pull(input_df_for_version, var2), 
                            plot = FALSE,
    lag.max=3)

  ccf_res <- tibble(lag = cross_correlations$lag[,,1], 
                  correlation=cross_correlations$acf[,,1])

  m <- ccf_res %>% filter(correlation == max(correlation))

  tibble(vlabel = unique(input_df_for_version$vlabel),
         fips = unique(input_df_for_version$state),
         max_lag = m$lag,
         max_corr = m$correlation,
         comparison = comparison,
         name = unique(input_df_for_version$name))

cross_corr <- state_res_differenced %>%
  group_by(state, vlabel) %>%
  group_split() %>%
  map_df(get_max_lag_by_version)

states_with_lag <- cross_corr %>%
  filter(max_lag >  0 & 
           vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$" & 
           max_corr > .55) %>% 
  pull(fips) %>%
  unique()

```{r, include=FALSE}

state_res_differenced %>%
  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
  ggplot(aes(x=exp_cases_median_differenced, y = infections_differenced)) +
  geom_point()

state_res %>%
  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
  ggplot(aes(x=exp_cases_median, y = infections)) +
  geom_point()

@@

## Compare States that Have a Lag

```{r,  fig.height=11, fig.width=12}

# states_with_lag

subset %>%
    filter(state %in% states_with_lag & biweek <= 25) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = 'Probabilistic Bias Analysis'),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  # geom_vline(xintercept=end_old_model,
  #            linetype=2,
  #            color="darkred",
  #            linewidth=.5) +
  facet_wrap(~name, ncol=3, scales = "free",
              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3.5 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=7),
          axis.text.y = element_text(size=4.5),
          legend.position = "top",
          legend.text=element_text(size=12),
          strip.text.x=element_text(size=11,
                                    face='plain',
                                    color='white')) +
  scale_fill_manual(values =comp_state_pal, name='')  +
  labs(y = "Estimated Infections",
       title = paste0("Estimated Infections by State\nin States with a Lag")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))

```{r, fig.height=11, fig.width=12, eval=save_figures}

ggsave(here('figures/intervals-and-covidestim-states-lag.pdf'))

## Test Positivity and Positive Tests in States with a Lag

```{r testpos-lag, fig.height=14, fig.width=12}

############################################
# POSITIVE TESTS AND TEST POSITIVITY
############################################
pltlist <- subset %>%
  filter(biweek<=25) %>%
  filter(state %in% states_with_lag) %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(state) %>%
  mutate(adj = (1/max(testpos)) * max(positive)) %>%
  group_split() %>%
  map(~ {
    adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=positive, color = 'Positive Tests')) +
      geom_point(aes(y=positive, color = 'Positive Tests'),
              alpha=.5,
              size=.8) +
    geom_line(aes(y=testpos*adj, color='Test Positivity'),
              linetype=2) +
      geom_point(aes(y=testpos*adj, color='Test Positivity'),
              alpha=.5,
              size=.8) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Test Positivity"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
                                                    linewidth=c(2,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=10),
          axis.text =element_text(size=9),
          legend.position="none") +
    scale_y_continuous(labels=comma, 
                       sec.axis = sec_axis(
                         ~./adj,
                         name = 'Test Positivity')) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name)

  })

legend <- get_legend(
  pltlist[[1]] +
    theme(legend.position = "top",
          legend.text=element_text(size=16))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)

```{r, eval=save_figures, fig.height=14, fig.width=12}

ggsave(here('figures/testpos-lag.pdf'))

### 2023-08-26 writing/supplement.Rmd
update manuscript & supplement
update figures

## COVID-19 Trends and Impact Survey

```{r, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$. Because of the noise present inthe data, we use the spline smoothed values to inform the priors. More specifically, we use cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021). While there are estimates of $\\Pr(S_1|\\text{untested})$ for all states (seen in the following figure), there are 27 states with estimates of $\\beta$.", fig.width=14,fig.height=12}
include_graphics(here('figures/smoothing-beta.pdf'))
```

@@

@@

```{r sim-intervals-faceted-county, fig.width=15, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts. Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval. The columns represent different implementations of the probabilistic bias analysis: in the first through the third columns, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered. For the implementation in the first column, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval. The second column centers only $\\beta$ at the aformentioned value, and the third column only centers $\\Pr(S_1|\\text{untested})$ at the aforementioned value. The fourth column corresponds to the implementation where we specify priors that are the same for all dates. The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date."}

>>>>

# COVID-19 Trends and Impact Survey

```{r, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$. Because of the noise present inthe data, we use the spline smoothed values to inform the priors. More specifically, we use cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021). While there are estimates of $\\Pr(S_1|\\text{untested})$ for all states (seen in the following figure), there are 27 states with estimates of $\\beta$.", fig.width=14,fig.height=12}

include_graphics(here('figures/smoothing-beta.pdf'))

@@

# Comparing Lag Between Probabilistic Bias Estimates and Covidestim Infections

```{r}

include_graphics(here('figures/intervals-and-covidestim-states-lag.pdf'))

```{r}

include_graphics(here('figures/testpos-lag.pdf'))

@@

```{r sim-intervals-faceted-county, fig.width=15.5, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts. Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval. The columns represent different implementations of the probabilistic bias analysis: in the first through the third columns, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered. For the implementation in the first column, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval. The second column centers only $\\beta$ at the aformentioned value, and the third column only centers $\\Pr(S_1|\\text{untested})$ at the aforementioned value. The fourth column corresponds to the implementation where we specify priors that are the same for all dates. The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date."}
